---
title: "Modern Statistical Computing"
subtitle: "Seminar 1"
author: "Martin Wiegand"
institute: "Pompeu Fabra University"
execute:
  echo: true
format:
  revealjs:
    theme: [default, custom.scss]
    scrollable: true
toc: true
toc-depth: 1
number-sections: false
mouse-wheel: true
code-overflow: scroll
code-line-numbers: false
code-copy: true
cache: true
title-slide-style: pandoc 
bibliography: references_seminar.bib
---

  
## Welcome! 
  
Brief organizational details of the seminars: 

- Will probably meet in weeks 2 (today), 4, 6, 8, 9 and 10 
- Idea of first seminars: you will apply the tools from class, I will assist you
- Some of the seminars towards the end will be devoted to your final projects 

Briefly about me: 

- 3rd-year PhD student in Econ 
- Contact: <martin.wiegand@upf.edu>
- In case of any questions/suggestions (also w.r.t your final projects), feel free to reach out! 

## Idea of Today's Seminar  

You will: 

1. pull some data from [Eurostat](https://ec.europa.eu/eurostat/data/database) on economic indicators and migrant settlement patterns across European regions

2. do some data manipulation, plots, etc. 

## Pull Eurostat Data 

There are various ways how we could pull data from Eurostat: 

**1.** Go to [Eurostat](https://ec.europa.eu/eurostat/data/database) website and manually select table/years/variables, etc.  
$\rightarrow$ Cons: Need to manually re-download data in case of changes to selection criteria (e.g. years, variables)  

**2.** For many publicly available databases, there are R packages available 
that allow you to skip 1. and directly import data to R.  
$\rightarrow$ Examples:  [Eurostat](https://ec.europa.eu/eurostat/data/database), [OECD](https://github.com/expersso/OECD), [World Bank](https://cran.r-project.org/web/packages/wbstats/vignettes/wbstats.html), $\dots$

In practice, often a combination of **1.** and **2.** (e.g. find table names of interest via **1.**, then download via **2.**)

--- 



## Prerequisites 

```{r, warning = FALSE}
library(eurostat)
# Alternatively, read-in data via readxl 
library(readxl)

# Data manipulation, etc 
library(tidyverse)
library(dplyr)
library(lubridate) # helpful for date manipulation 
```

```{r}
migrant_df <- get_eurostat("cens_11cobe_r2")

migrant_df

```

## Homework 

In the seminar sessions that are not devoted to your final projects, parts of the assignment will be considered as homework, and are **due on the next Thursday (one week later) at 3pm**. 

**Preferred format** of handed in homework: rendered R-markdown or quarto notebook (as PDF or HTML).  










  
  
  
    